---
title: "p8105_hw1_wc2788"
author: "Wen Cheng"
date: "9/29/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Problem 1
### Create a data frame comprised of:    

1. a random sample of size 10 from a standard Normal distribution
2. a logical vector indicating whether elements of the sample are greater than 0
3. a character vector of length 10
4. a factor vector of length 10, with 3 different factor “levels"

```{r }
dataframe = tibble(
  random_sample = rnorm(10),
  logical_vec = random_sample > 0,
  char_vec =  c("a","b","c","d","e","f","j","h","i","j"),
  factor_vec = factor(c("high","medium","low","high","medium","low","high","medium","low","high") )
)
```

### Take the mean of each variable
```{r}
#use pull() to pull the variable out of the dataframe
#random sample / numbers vector
mean(pull(dataframe,random_sample))
#logical vector
mean(pull(dataframe,logical_vec))
#character vector
mean(pull(dataframe,char_vec))
#factor vector
mean(pull(dataframe,factor_vec))
```
can take the mean of numbers and logical but not character or factor

### Applies the as.numeric function to the logical, character and factor
```{r, eval=FALSE}
#logical vector
as.numeric(pull(dataframe,logical_vec))
#character vector
as.numeric(pull(dataframe,char_vec))
#factor vector
as.numeric(pull(dataframe,factor_vec))
```
Logical and factor vector can convert to numerical and take the mean, however, character vector cannot.   

The reason of these is that    
1. in logical vector, TURE can convert to 1 and FALSE can convert to 0    
2. in factor vector, there are 3 levels. When we use as.numeric(), the default sorting of the levels of the factors is sorted in alphabetical order. We can find that it converted "high" to 1, "low" to 2, and "medium" to 3     
3. in character vector, it contains several strings. strings cannot simply convert to numbers   

It also explained why logical vector has a mean, but factor and character not. In the definition, mean(x, trim = 0, na.rm = FALSE, …), x can be numeric/logical vectors and date, date-time and time interval objects. 


## Problem 2
### Load the penguins dataset
```{r}
data("penguins", package = "palmerpenguins")
print(penguins)
```
### Short description of the penguins dataset
The penguins dataset contains penguin species (`r paste(levels(penguins$species), collapse = ", ")`), islands( `r paste(levels(penguins$island), collapse = ", ")`) and bill length, bill depth, flipper length, body mass, sex, and year.   
The dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns.   
The mean flipper length is `r mean(penguins$flipper_length_mm,na.rm=TRUE)` mm. 

### Scatterplot
```{r, warning=FALSE}
#make a scatterplot
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()
#export the scatterplot
ggsave("first_scatterplot.pdf")
```
